---
title: "NE GA Vaccine and Service Providers"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: lumen
---

```{r setup, include=FALSE}

#load required packages
library(flexdashboard)
library(sf)
library(shiny)
library(leaflet)
library(BAMMtools) #nice jenks

library(dplyr)

#load and format data #####

#this does not need to be run everytime but is part of the workflow to download
#source("source/download_census_vars.R")

#this cleans and subsets the spatial variables
#it also defines the breaks for each individual variable
source("source/format_spatial_vars.R", local = knitr::knit_global())

```

Service Area Maps
======================================

Column {.sidebar}
-----------------------------------------------------------------------

### Map of demographics, service providers, and vaccine providers

Click on a provider point for more information.

```{r}
selectInput("county_choice", "Select a county:",
            choices = c("All", "Barrow", "Clarke", "Elbert", "Greene", "Jackson",
                        "Madison", "Morgan", "Oconee", "Oglethorpe", "Walton"),
            selected = "All")
```



```{r}
selectInput("census_choice", "Select a census variable:",
            choices = c("Civilians with Disabilities" = "disable_pop", 
                        "Households without Vehicles" = "noveh_pop",
                        "Minority Population" = "nonwht_pop",
                        "Persons (5+) Who Speak English Less Than Well" = "engnotwell_pop",
                        "Persons Over 65 Years Old"= "age65_pop", 
                        "Persons Below Poverty" = "pov_pop", 
                        "Persons (25+) without High School Diploma" = "lesshsdip_pop"),
            selected = "pov_pop")
```


```{r}
sliderInput("poly_alpha", "Census Data Opacity:",
            min = 0, max = 1, value = 0.5)
```

Designed by Michelle Evans & Jerry Shannon at the [Community Mapping Lab](http://communitymappinglab.org)

Column
-----------------------------------------------------------------------


```{r}
output$map1 <- renderLeaflet({
  
  #select variable
  census.map <- select(census.vars, County, input$census_choice, popup_text) 
  colnames(census.map)[2] <- "value"
  
  #create color palette and labels
  census.cols <- data.frame(color_pal = c("YlGnBu", "Reds",  "Purples", 
                   "Blues", "Oranges", "Greens", "PuRd"),
                   census_var = c("disable_pop", "noveh_pop", "nonwht_pop", 
                                  "engnotwell_pop", "age65_pop", "pov_pop", "lesshsdip_pop"),
                   census_label = c("Civilians with Disabilities",
                                    "Households without Vehicles",
                                    "Minority Population",
                                    "Persons (+5) Who Speak\nEnglish Less Than Well",
                                    "Persons Over 65 Years Old",
                                    "Persons Below Poverty",
                                    "Persons (25+) without\nHigh School Diploma")) %>%
    mutate(color_pal=as.character(color_pal),
           census_label=as.character(census_label))
  
  current.setting <- filter(census.cols, census_var == input$census_choice)
  legend_title<-current.setting$census_label

  #need range for domain
  col.domain <- c(0,max(census.map$value))
  col.breaks <- getJenksBreaks(c(0,census.map$value), 6, subset = NULL)

  binpal <- colorBin(
  palette = current.setting$color_pal,
  domain = col.domain,
  bins = col.breaks
  )
  
  if(input$county_choice != "All"){
    census.map <- filter(census.map, County == input$county_choice)
  }

  county_map<-service.area
  if(input$county_choice != "All"){  
    county_map<-service.area %>% filter(NAME==input$county_choice)
  }
  
#crop points to a 5km buffer around this area
  crop.points <- services.all[st_buffer(census.map, 0.05),]

  map1 <- leaflet(census.map) %>%
    addTiles(urlTemplate="https://{s}.basemaps.cartocdn.com/light_all/{z}/{x}/{y}{r}.png",
             attribution='&copy; <a href="https://www.openstreetmap.org/copyright">OpenStreetMap</a> contributors &copy; <a href="https://carto.com/attributions">CARTO</a>') %>%
        addPolygons(data=county_map,
                fillOpacity=0,
                color="black",weight=1) %>%
    addPolygons(data=census.map,
                fillOpacity = input$poly_alpha,
              fillColor = ~binpal(value),
              stroke = T,
              color = "black", opacity = 0.4, weight = 0.5,
              popup = ~popup_text) %>%
    addCircles(data = crop.points, color = "black", opacity = 1,
             #popup = ~textbox, 
             layerId = ~prov_id) %>%
    addLegend("bottomright",
              pal = binpal,
              values = ~value,
              title = current.setting$census_label,
              opacity = 1)

  map1
})

leafletOutput('map1', height = "95vh")
```


Column {data-width=250}
-----------------------------------------------------------------------

### Provider Info

```{r}

observe({
  validate(need(input$map1_shape_click, FALSE))
  validate(need(input$map1_shape_click %in% services.all$prov_id,FALSE))

  serv_text <- services.all %>% filter(prov_id==input$map1_shape_click$id)
  output$text1 <- renderUI({
    HTML(serv_text$textbox)
  })
})

htmlOutput("text1",width=200)
```


About this map
=====================

# Purpose:
This dashboard provides an interactive visualization of COVID-19 vaccinations sites and seven demographic characteristics for census tracts in the following counties of Northeast Georgia: **Barrow**, Clarke**, **Elbert**, Greene**, Jackson**, **Madison**, **Morgan, Oconee**, **Oglethorpe**, and **Walton**. 
## How to use this map:
Begin by selecting the county of interest from the first drop down menu. Then select the census variable you would like to be shown. Use the opacity slider to adjust the transparency of the colors – a higher opacity will make the colors appear darker. Click on a census tract or vaccination site to view detailed information in a popup window. **Note: when viewing data for a single county you will see all vaccination providers inside the county and within 5km from the county boundary.** 
### Data Sources: 
The demographic data is from the 2015-2019 American Community Survey and the vaccination site data was produced by the Department of Public Health.    
